### 2023-06-30 paper/main.tex
corrections_applied

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage[english]{babel}          % für deutsche Dokumente
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{overpic,subfigure, tikz}

@@

\title{Local Wave Field Synthesis by Temporal Bandlimitation*
    \thanks{This work was supported by the János Bolyai Research Scholarship of the Hungarian Academy of Sciences, the ÚNKP-22-5-BME-318 New National Excellence Program of the Ministry for Innovation and Technology from the source of the National Research, Development and Innovation Fund
        and by the OTKA PD-143129 and OTKA K-143436 grants.
    }
}

@@

As WFS obtains the necessary driving functions from a suitable boundary integral representation of the target sound field, perfect phase correct field-reconstruction could be only achieved by employing a theoretical, continuous secondary source distribution.
In practical applications the SSD consist of discrete loudspeaker elements, resulting in secondary wavefronts additionally to the intended virtual wavefront, emerging from the individual SSD elements, termed as aliasing waves.
This spatial aliasing phenomena is perceived as the coloration of the synthesized sound field above the spatial aliasing frequency, varying with the listening position and the virtual wavefront direction, degrading the quality of reproduction \cite{8371275, 8611109}.

@@

The paper is organized as follows:
Section \ref{sec:theory} gives a brief theoretical introduction on basic WFS theory and spatial aliasing phenomena. 
Section \ref{sec:novel_approach} first discusses the spatial-to-temporal filter transform, then introduces the novel local WFS methodology.
Section \ref{sec:evaluation} gives a comparison of the presented technique with a recent LWFS approach based on direct modal bandlimitation, followed by a short summary and outlook in section \ref{sec:conclusion}.

@@

\subsection{The local propagation vector:}
Consider an arbitrary steady-state free space sound field at an angular frequency $\omega$.
As a standard ansatz in the field of geometrical acoustics the sound field is written in the general polar from as
%
\begin{equation}
    P(\vx,\omega) = A(\vx) \te^{- \ti \omega \tau(\vx)},
    \label{eq:Pxw}
\end{equation}
%
where $A(\vx)$ and $\tau(\vx)$ are real-valued functions.
This formulation applies to both plane waves and (3D) point sources.
The propagation dynamics of the sound field are governed by its phase function $\tau(\vx)$, termed as the eikonal in the field of ray acoustics \cite{Bleistein2000}.
In the temporal domain, the sound field can be obtained by taking the inverse Fourier transform of \eqref{eq:Pxw}, yielding:
\begin{equation}
    p(\vx,t) = A(\vx) \, \delta \! \left( t - \tau(\vx)\right).
\end{equation}
From the above formulation it is evident that the eikonal $\tau(\vx)$ describes the propagation delay the wavefront takes to arrive at $\vx$.
The eikonal equation is obtained by substituting the ansatz \eqref{eq:Pxw} into the Helmholtz equation, stating that in a source free volume
\begin{equation}
    \left| \nabla \tau(\vx) \right| = \frac{1}{c}
\end{equation}
is fulfilled.
The gradient of the eikonal is termed as the local propagation vector
\begin{equation}
    \hat{\vk}^P(\vx) = \posvec{3}{\hat{k}_x^P(\vx)}{\hat{k}_y^P(\vx)}{\hat{k}_z^P(\vx)} = c \nabla \tau(\vx,\omega).
\end{equation}
In steady-state, the local propagation vector is perpendicular to the wavefront (equiphase positions) at an arbitrary location of unit length pointing towards the local propagation direction \cite{Firtha2016}.
In the temporal domain it points into the local propagation direction at the time instant of the wavefront passby.

@@

\subsection{Conventional 2.5D Wave Field Synthesis}
Consider a smooth convex secondary source distribution located at $\vxo = \posvec{3}{x_0}{y_0}{0}$ consisting of a continuous distribution of 3D point sources, described by the 3D freefield Green's function, reading in the frequency domain as
\begin{equation}
    G_{3 \mathrm{D}}(\vx,\vxo,\omega) = \frac{1}{4\pi}\frac{\te^{-\frac{\omega}{c}\left| \vx-\vxo\right|}}{\left| \vx-\vxo\right|},
\end{equation}
where $\omega$ denotes the angular frequency.
In this geometry the target field inside the area bounded by the SSD is described by the  Kirchhoff approximation of 2.5D Kirchhoff-Helmholtz integral, from which the 2.5D driving functions can be extracted.
For an arbitrary simple target virtual sound field $P(\vxo,\omega)$ the steady-state driving functions read as \cite{Firtha2016}
\begin{equation}
    D(\vxo,\omega) = \underbrace{\sqrt{8\pi \frac{\ti \omega}{c}}}_{ H_{\text{pre}}(\omega) }
    w(\vxo) \sqrt{d_{\mathrm{ref}}(\vxo)}
    P(\vxo,\omega)
    .
    \label{Eq:WFS25D}
\end{equation}
The driving function consists of
\begin{itemize}
    \item a frequency dependent pre-equalization filter $H_{\text{pre}}(\omega)$ \cite{Start1997:phd, Berkhout1993:Acoustic_control_by_WFS, Bleistein1986},
    \item a secondary source selection window \cite{Nicol1999, Spors2007}
          \begin{equation}
              w(\vxo) = \max{\left( \hat{k}_{\mathrm{n}}^P(\vxo),0\right)},
          \end{equation}
          where $\hat{k}_{\mathrm{n}}^P(\vxo)$ is the normal component of the local propagation vector,
    \item a gain factor $\sqrt{d_{\mathrm{ref}}(\vxo)}$ allowing amplitude correction along a reference curve, depending on the actual virtual field model (c.f. \cite{Start1997:phd} and for a more complete treatise \cite{Firtha2016}),
    \item and the virtual field measured on the SSD.
\end{itemize}
Assuming a simple virtual sound field as given by \eqref{eq:Pxw} the driving functions can be written as
\begin{equation}
    D(\vxo,\omega) = H_{\text{pre}}(\omega)
    \underbrace{ w(\vxo)  \, \sqrt{d_{\mathrm{ref}}(\vxo)} \, A(\vxo) }_{A^D(\vxo)} \te^{- \ti \omega \tau(\vxo)}
    ,
\end{equation}
and in the time domain as
\begin{equation}
    d(\vxo,t) = h_{\text{pre}}(t) \ast_t  A^D(\vxo)\,  \delta\left( t - \tau(\vxo)\right),
    \label{eq:d_wfs_td}
\end{equation}
with $A^D(\vxo)$ being the real valued overall gain factor of the driving function, $\ast_t$ denoting temporal convolution and $h_{\text{pre}}(t)$ is the temporal WFS prefilter impulse response.
The analytical, continuous pre-equalization filter impulse response is given in \cite{Deregowski1983, Firtha2019phd}, while for an ideal, non-casual FIR filter implementation the reader is referred to \cite{Schultz2016}.
In the following due to the associativity of convolution this prefiltering is excluded from the discussion.

@@

\begin{figure*}[h!]
    \begin{center}
        \begin{overpic}[width = 0.75\columnwidth]{figs/ideal_synthesis.png}
            \footnotesize \put(0,2){(a)}
        \end{overpic} \hspace{2cm}
        \begin{overpic}[width = 0.75\columnwidth]{figs/aliased_synthesis.png}
            \footnotesize \put(0,2){(b)}
        \end{overpic}
        \\
        \begin{overpic}[width = 0.75\columnwidth]{figs/ideal_synthesis_spectrum.png}
            \footnotesize \put(0,0){(c)}
        \end{overpic} \hspace{2cm}
        \begin{overpic}[width = 0.75\columnwidth]{figs/aliased_synthesis_spectrum.png}
            \footnotesize \put(0,0){(d)}
        \end{overpic}
    \end{center}
    \caption{The synthesized field (top) and the spatial spectrum (bottom) of the driving function in case of ideal, quasi-continuous SSD (a) and (c), and in case of a discretized SSD (b) and (d).
    It is important to note that the spatial spectrum depicts only the non-zero Fourier components (i.e. the Fourier series coefficients) normalized to the sampling wavenumber.}
    \label{Fig:aliasing}
\end{figure*}
The backbone of WFS theory assumes a continuous secondary source distribution.
In practical applications the SSD is composed of evenly spaced discrete source elements.
As a result, \emph{aliasing wavefronts} emerge from the individual secondary sources, following the intended virtual wavefront, leading to a \emph{spatial aliasing phenomena}.
Although localization does not degrade due to the precedence effect, aliasing is perceived as strong coloration predominantly in the high-frequency region varying with the receiver position and the virtual source position/direction, occurring predominantly above the \emph{aliasing frequency}.

Figure \ref{Fig:aliasing} illustrates spatial aliasing in the temporal domain through the example of the synthesis of an impulsive plane wave, arriving from $\phi_{\mathrm{PW}} = 0^{\circ}$, applying a discrete secondary source distribution in comparison with a quasi-continuous synthesis scenario.
In the current example the SSD is a circular one with the radius of $R_0 = 2~\mathrm{m}$, consisting of $N_0 = 90$ 3D point sources.
The aliased synthesis in the temporal domain is depicted in Figure \ref{Fig:aliasing} (b).

Mathematically, aliasing can be modeled as the discretization of the theoretically continuous driving functions:
Assume that a suitable parametrization $d(s,t)$ of the secondary source is given, e.g. the arc length $s = R_0\, \cos \phi$ in case of a circular SSD (where $\phi$ is the polar angle), or the linear position on an infinite long linear SSD.
The discretization process is modeled by sampling the driving function by a sampling function, consisting of a series of Dirac-deltas, where the impulse occurences match the actual SSD positions:
\begin{equation}
    d_s(s,t) = \sum_{\nu = 0}^{N_0-1} d(s,t) \, \delta(s - \nu \Delta s) \, \Delta s,
    \label{eq:sampling_td}
\end{equation}
where $\Delta s$ is the sampling arc length.

By investigating the sampling process in the frequency-wavenumber domain, the spectrum of the sampled driving function is given by
\begin{equation}
    \tilde{D}_s(k_s,\omega) = \sum_{\nu = -\infty}^{\infty} \tilde{D}\left(k_s- \nu \frac{2\pi}{\Delta s},\omega\right) \Delta s,
    \label{eq:sampling_fd}
\end{equation}
where $t \rightarrow \omega$ and $s \rightarrow k_s$ are variables pairs for the Fourier-transform.
Obviously, for a closed SSD contour (most notably for a circular SSD) the continuous spatial Fourier-transform yields a discrete spectrum, consisting of non-zero components only on the multiples of the spatial wavenumber, i.e. at
\begin{equation}
    k_s = \frac{m}{R_0},
\end{equation}
constituting a Fourier series.

In the followings the spatial spectrum refers to the distribution of the non-zero wavenumber values, and the corresponding wavenumber refers to the spatial frequency (multipled by $2\pi$) of the harmonic basis functions, measured along the closed curve.
Figure \ref{Fig:aliasing} (c) and (d) illustrate the quasi-continuous and the sampled driving function's spectra.

From \eqref{eq:sampling_fd} it can be concluded that due to spatial sampling, the wavenumber content of the driving function is repeated on the multiples of the sampling wavenumber
\begin{equation}
k_{s, \mathrm{s}} = \frac{2\pi}{\Delta s}.
\end{equation}
Since the conventional driving functions are not-bandlimited, therefore, the repeating spectra will overlap above the Nyquist wavenumber $k_{s, \mathrm{Nyq}} = \frac{\pi}{\Delta s}$, for which the corresponding angular frequency is the aliasing frequency of the discretization scheme, as depicted in Figure \ref{Fig:aliasing} (d).

According to the sampling theorem, spatial aliasing can be avoided by analytically spatial-bandlimiting the driving functions to the Nyquist wavenumber.
This can be performed by either spatially filtering the driving functions with an appropriate spatial low-pass filter prior to numerical evaluation, or performing bandlimitation directly in the wavenumber (modal) domain analytically.
The latter approach is implemented by \cite{hahn2022cylindrical}, and will serve as a reference solution in the following investigation.
However, the modal bandlimitation solution is available only for plane wave and point source virtual source models synthesized by a circular SSD, at the cost of increased computational complexity.

In the following an alternative, approximate solution is presented, allowing spatial antialiasing filtering for an arbitrary virtual source model and arbitrary SSD contour, performed by FIR filtering in the temporal domain.
The technique relies on an analytic transform of spatially defined antialiasing filters to an equivalent temporal domain filter bank.

@@

\subsection{Spatial to temporal filter transform}
First, a general analytical approach is introduced allowing the transform of an arbitrary spatial filter to an equivalent temporal filter set.
From herein, subscript $x$ and $t$ distinguishes wether a filter is defined as a spatial or a temporal filter, respectively.

Assume an arbitrary filter impulse response defined in the spatial domain, denoted by $h_x(s)$.
Again, $s$ is a suitable parametrization of the SSD contour.
The spatially filtered driving function is obtained from the convolution of the conventional driving functions and the filter impulse response along the SSD contour
\begin{equation}
    d'_x(s,t) = h_x(s) \ast_s d(s,t) = \int h_x(s-s_0) \, d(s_0,t) \, \td s_0,
    \label{eq:spatial_conv}
\end{equation}
where $\ast_x$ denotes a circular convolution for a convex SSD contour or a linear convolution in case of an infinite long linear SSD.

@@

\subsubsection{The exact transformation}
The following transformation is allowed by the spatio-temporal structures of the conventional driving functions \eqref{eq:d_wfs_td}, in which the spatial and temporal dimensions are interconnected in the argument of a Dirac-delta, due to the wave propagation characteristics.
The main idea of the derivation is that by exploiting the sifting property of the Dirac-delta in the driving functions, both spatial and temporal convolutions \eqref{eq:spatial_conv} and \eqref{eq:temp_conv} can be evaluated, from which the temporal filter impulse responses can be expressed as the function of the spatial ones.
Again, in the following the pre-equalization filter is omitted from discussion, since equalization filtering can be performed following the antialiasing process.

The temporal convolution in \eqref{eq:temp_conv} can be evaluated by substituting the general WFS driving function \eqref{eq:d_wfs_td} and exploiting the sifting property of the Dirac-delta, yielding
\begin{align}
    d'_t(s,t) = A^D(s) \, h_t \!\left(s,t- \tau(s) \right) \\
    d'_t(s,t + \tau(s) ) = A^D(s) \, h_t \!\left(s,t\right).
    \label{eq:temp_conv_eval}
\end{align}
In the second equation the time shift of $\tau(s)$ was applied for the sake of simplicity.

The shifted spatially filtered driving function reads as
\begin{multline}
    d'_x(s,t+ \tau(s) ) = \int h_x(s-s_0) \, d(s_0,t+ \tau(s) ) \, \td s_0 = \\
    \int h_x(s-s_0) \, A^D(s_0)\, \delta(s_0,t+ \left(\tau(s)-\tau(s_0)\right)  ) \, \td s_0.
    \label{eq:spat_conv2}
\end{multline}
In order to evaluate the spatial convolution the generalized sifting property of the Dirac-delta may be applied \cite{Firtha2019phd, Jackson1999}, which states that
\begin{equation}
    \int f(s_0) \delta(g(s_0)) \td s_0 = \sum_i \frac{f(s_i)}{| \frac{\partial}{\partial s_0} g(s_0) |_{s = s_i}}, \hspace{3mm} g(s_i) = 0.
\end{equation}
In the present problem the zeros of the Dirac's argument are found where
\begin{equation}
    \tau(s_i) = t + \tau(s)
    \label{eq:zero}
\end{equation}
is satisfied.
In the following it is assumed that a single zero $s_i$ exists, satisfying \eqref{eq:zero}.
This assumption means that the wavefront arrives at each SSD element at a unique time instant, being strictly true for a virtual plane wave and a linear SSD.
The derivative of the phase function is given by
\begin{equation}
    \frac{\partial}{\partial s_0} g(s_0) = - \frac{\partial}{\partial s_0} \tau(s_0) = -\frac{\hat{k}_t^P(s_0)}{c},
\end{equation}
i.e. it is the tangential component of the local propagation vector.
With the above assumptions the integral \eqref{eq:spat_conv2} can be evaluated, yielding
\begin{equation}
    d'_x(s,t+ \tau(s) ) =  \frac{ c }{|\hat{k}_t^P(s_i)|} \, h_x(s-s_i) \, A^D(s_i)
    .
    \label{eq:spat_conv_eval}
\end{equation}

@@

Equation \eqref{eq:temporal_transform} can be directly expressed in the spectral domain by taking the Fourier transform of both sides, relating the frequency response of the spatial and the temporal filters.
By denoting $\mathcal{F}\left( h_x(s)\right) = H_x(k_s)$ the corresponding transform is given by
\begin{equation}
    H_t\left(s,\omega\right) = H_x\left(-\frac{\omega}{c}\hat{k}_t^P(s)\right) = H_x\left(-k_t^P(s)\right),
    \label{eq:frequency_transform}
\end{equation}
where $k_t^P(s)$ is the tangential component of the local wavenumber vector \footnote{The local wavenumber vector can be defined for a steady state sound field, pointing in the local propagation direction with the length being $\omega/c$ \cite{Firtha2019phd}}.
Hence, as the main result of the present discussion, the temporal filter transfer can be obtained from the wavenumber (modal) content of the spatial filter by simple rescaling in terms of the local wavenumber vector.

@@

\subsubsection{Application for Local Wave Field Synthesis}
The findings of the previous subsections can be straightforwardly used in order to apply the above theory directly for Local Wave Field Synthesis.
In this case a reference position $\vxref$ is prescribed inside the SSD, at which aliasing-free, amplitude and phase correct synthesis has to be ensured.

>>>>

\documentclass[conference]{IEEEtran}
\IEEEoverridecommandlockouts
% The preceding line is only needed to identify funding in the first footnote. If that is unneeded, please comment it out.
\usepackage[english]{babel}          % für deutsche Dokumente
\usepackage{cite}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{algorithmic}
\usepackage{graphicx}
\usepackage{textcomp}
\usepackage{xcolor}
\usepackage{nicefrac}
\usepackage{overpic,subfigure, tikz}

@@

\title{Local Wave Field Synthesis by Temporal Bandlimitation
    \thanks{This work was supported by the János Bolyai Research Scholarship of the Hungarian Academy of Sciences, the ÚNKP-22-5-BME-318 New National Excellence Program of the Ministry for Innovation and Technology from the source of the National Research, Development and Innovation Fund
        and by the OTKA PD-143129 and OTKA K-143436 grants.
    }
}

@@

As WFS obtains the necessary driving functions from a suitable boundary integral representation of the target sound field, perfect phase correct field-reconstruction could be only achieved by employing a theoretical, continuous secondary source distribution.
In practical applications the SSD consist of discrete loudspeaker elements, resulting in secondary wavefronts additionally to the intended virtual wavefront, emerging from the individual SSD elements, termed as aliasing waves.
This spatial aliasing phenomenon is perceived as coloration of the synthesized sound field above the spatial aliasing frequency, varying with the listening position and the virtual wavefront direction, degrading the quality of reproduction \cite{8371275, 8611109}.

@@

The paper is organized as follows:
Section \ref{sec:theory} gives a brief theoretical introduction on basic WFS theory and spatial aliasing phenomena. 
Section \ref{sec:novel_approach} first discusses the spatial-to-temporal filter transform, then introduces the novel local WFS methodology.
Section \ref{sec:evaluation} gives a comparison of the presented technique with a recent LWFS approach based on direct modal bandlimitation, followed by a short summary and outlook in Section \ref{sec:conclusion}.

@@

\subsection{Local propagation vector concept}
Consider an arbitrary steady-state free space sound field at an angular frequency $\omega$.
As a standard ansatz in the field of geometrical acoustics the sound field at a spatial position $\vx = [x\, y\, z]^{\mathrm{T}}$ is written in the general polar from as
%
\begin{equation}
    P(\vx,\omega) = A(\vx) \te^{- \ti \omega \tau(\vx)},
    \label{eq:Pxw}
\end{equation}
%
where $A(\vx)$ and $\tau(\vx)$ are real-valued functions.
This formulation applies to both plane waves and (3D) point sources.
The propagation dynamics of the sound field are governed by its phase function $\tau(\vx)$, termed as the eikonal in the field of ray acoustics \cite{Bleistein2000}.
In the temporal domain, the sound field can be obtained by taking the inverse temporal Fourier-transform of \eqref{eq:Pxw}, yielding
\begin{equation}
    p(\vx,t) = A(\vx) \, \delta \! \left( t - \tau(\vx)\right).
\end{equation}
From the above formulation it is evident that the eikonal $\tau(\vx)$ describes the propagation delay, the wavefront takes to arrive at $\vx$.
The eikonal equation is obtained by substituting the ansatz \eqref{eq:Pxw} into the Helmholtz equation, stating that in a source free volume
\begin{equation}
    \left| \nabla \tau(\vx) \right| = \frac{1}{c}
\end{equation}
is fulfilled.
The gradient of the eikonal weighted by the speed of sound is termed as the local propagation vector
\begin{equation}
    c \nabla \tau(\vx,\omega) = \hat{\vk}^P(\vx) = \posvec{3}{\hat{k}_x^P(\vx)}{\hat{k}_y^P(\vx)}{\hat{k}_z^P(\vx)}.
\end{equation}
In steady-state, the local propagation vector is perpendicular to the wavefront (equiphase positions) at an arbitrary location of unit length pointing towards the local propagation direction \cite{Firtha2016}.
In the temporal domain it points into the local propagation direction at the time instant of the wavefront passby.

@@

\subsection{2.5D Wave Field Synthesis}
Consider a smooth convex secondary source distribution located at $\vxo = \posvec{3}{x_0}{y_0}{0}$ consisting of a continuous distribution of 3D point sources, described by the 3D freefield Green's function, reading in the frequency domain as
\begin{equation}
    G_{3 \mathrm{D}}(\vx,\vxo,\omega) = \frac{1}{4\pi}\frac{\te^{-\ti \frac{\omega}{c}\parallel \vx-\vxo \parallel }}{\parallel \vx-\vxo\parallel },
\end{equation}
where $\omega$ denotes the angular frequency.
In this geometry the target field inside the area bounded by the SSD is described by the  Kirchhoff approximation of 2.5D Kirchhoff-Helmholtz integral, from which the 2.5D driving functions can be extracted \cite{Bleistein1986}.
For an arbitrary, simple target virtual sound field $P(\vxo,\omega)$ the steady-state driving functions read as \cite{Firtha2016}
\begin{equation}
    D(\vxo,\omega) = \underbrace{\sqrt{8\pi \frac{\ti \omega}{c}}}_{ H_{\text{pre}}(\omega) }
    w(\vxo) \sqrt{d_{\mathrm{ref}}(\vxo)}
    P(\vxo,\omega)
    .
    \label{Eq:WFS25D}
\end{equation}
The driving function consists of
\begin{itemize}
    \item a frequency dependent pre-equalization filter $H_{\text{pre}}(\omega)$ \cite{Start1997:phd, Berkhout1993:Acoustic_control_by_WFS, Bleistein1986},
    \item a secondary source selection window \cite{Nicol1999, Spors2007}
          \begin{equation}
              w(\vxo) = \max{\left( \hat{k}_{\mathrm{n}}^P(\vxo),0\right)},
          \end{equation}
          where $\hat{k}_{\mathrm{n}}^P(\vxo)$ is the normal component of the local propagation vector (i.e. its projection to the SSD normal vector at $\vxo$),
    \item a gain factor $\sqrt{d_{\mathrm{ref}}(\vxo)}$ allowing amplitude correction along a reference curve, depending on the actual virtual field model (cf. \cite{Start1997:phd, Firtha2016}),
    \item and the virtual field measured on the SSD.
\end{itemize}
Assuming a simple virtual sound field as given by \eqref{eq:Pxw} the driving functions can be written as
\begin{equation}
    D(\vxo,\omega) = H_{\text{pre}}(\omega)
    \underbrace{ w(\vxo)  \, \sqrt{d_{\mathrm{ref}}(\vxo)} \, A(\vxo) }_{A^D(\vxo)} \te^{- \ti \omega \tau(\vxo)}
    ,
\end{equation}
and in the time domain as
\begin{equation}
    d(\vxo,t) = h_{\text{pre}}(t) \ast_t  A^D(\vxo)\,  \delta\left( t - \tau(\vxo)\right),
    \label{eq:d_wfs_td}
\end{equation}
with $A^D(\vxo)$ being the real valued overall gain factor of the driving function, $\ast_t$ denoting temporal convolution and $h_{\text{pre}}(t)$ is the temporal WFS prefilter impulse response.
The analytical, continuous pre-equalization filter impulse response is given in \cite{Deregowski1983, Firtha2019phd}, while for an ideal, non-casual FIR filter implementation the reader is referred to \cite{Schultz2016}.
In the following due to the associativity of convolution this prefiltering is excluded from a more convenient discussion.

@@

\begin{figure*}[h!]
    \begin{center}
        \begin{overpic}[width = 0.75\columnwidth]{figs/ideal_synthesis.png}
            \footnotesize \put(0,2){(a)}
        \end{overpic} \hspace{2cm}
        \begin{overpic}[width = 0.75\columnwidth]{figs/aliased_synthesis.png}
            \footnotesize \put(0,2){(b)}
        \end{overpic}
        \\
        \begin{overpic}[width = 0.75\columnwidth]{figs/ideal_synthesis_spectrum.png}
            \footnotesize \put(0,0){(c)}
        \end{overpic} \hspace{2cm}
        \begin{overpic}[width = 0.75\columnwidth]{figs/aliased_synthesis_spectrum.png}
            \footnotesize \put(0,0){(d)}
        \end{overpic}
    \end{center}
    \caption{The synthesized field (top) and the spatial spectrum (bottom) of the driving function in case of ideal, quasi-continuous SSD (a) and (c), and in case of a discretized SSD (b) and (d).
    It is important to note that the spatial spectrum depicts only the non-zero Fourier components (i.e. the Fourier-series coefficients) normalized to the sampling wavenumber.}
    \label{Fig:aliasing}
\end{figure*}
The backbone of WFS theory assumes a continuous secondary source distribution.
In practical applications the SSD is composed of evenly spaced discrete source elements.
As a result, aliasing wavefronts emerge from the individual secondary sources, following the intended virtual wavefront, leading to spatial aliasing phenomena.
Although localization does not degrade due to the precedence effect, aliasing is perceived as strong coloration predominantly in the high-frequency region varying with the receiver position and the virtual source position/direction, occurring predominantly above the aliasing frequency.

Figure \ref{Fig:aliasing} illustrates spatial aliasing in the temporal domain through the example of the synthesis of an impulsive plane wave, arriving from $\phi_{\mathrm{PW}} = 0^{\circ}$, applying a discrete secondary source distribution in comparison with a quasi-continuous synthesis scenario.
In the current example a circular SSD is applied with the radius of $R_0 = 2~\mathrm{m}$, consisting of $N_0 = 90$ 3D point sources.
The aliased synthesis in the temporal domain is depicted in Figure \ref{Fig:aliasing} (b).

Mathematically, aliasing can be modeled as the discretization of the theoretically continuous driving functions.
Assume that a suitable parametrization $d(s,t)$ of the secondary source is given, e.g. the arc length $s = R_0\, \cos \phi$ in case of a circular SSD (where $\phi$ is the polar angle), or the linear position on an infinite, linear SSD.
The discretization process is modeled by sampling the driving function by a sampling function, consisting of a series of Dirac-deltas, where the impulse occurences match the actual SSD positions:
\begin{equation}
    d_s(s,t) = \sum_{\nu = 0}^{N_0-1} d(s,t) \, \delta(s - \nu \Delta s) \, \Delta s,
    \label{eq:sampling_td}
\end{equation}
with $\Delta s$ being the sampling (arc) length.

By investigating the sampling process in the frequency-wavenumber domain, the spectrum of the sampled driving function is given by
\begin{equation}
    \tilde{D}_s(k_s,\omega) = \sum_{\nu = -\infty}^{\infty} \tilde{D}\left(k_s- \nu \frac{2\pi}{\Delta s},\omega\right),
    \label{eq:sampling_fd}
\end{equation}
where $t \rightarrow \omega$ and $s \rightarrow k_s$ are variable pairs for the Fourier-transform.
Obviously, for a closed SSD contour (most notably for a circular SSD) the continuous spatial Fourier-transform yields non-zero components at discrete spatial frequencies on the multiples of the spatial wavenumber, i.e. at
\begin{equation}
    k_s = \frac{m}{R_0},
\end{equation}
with $m \in \mathbb{Z}$ constituting a Fourier-series.

In the followings the spatial spectrum refers to the distribution of the non-zero wavenumber values, and the corresponding wavenumber refers to the spatial frequency (multiplied by $2\pi$) of the harmonic basis functions, measured along the closed curve.
Figure \ref{Fig:aliasing} (c) and (d) illustrate the quasi-continuous and the sampled driving function's spectra.

From \eqref{eq:sampling_fd} it can be deduced that due to spatial sampling, the wavenumber content of the driving function is repeated on the multiples of the sampling wavenumber
\begin{equation}
k_{s, \mathrm{s}} = \frac{2\pi}{\Delta s}.
\end{equation}
Since the conventional driving functions are not-bandlimited, therefore, the repeating spectra will overlap above the Nyquist wavenumber $k_{s, \mathrm{Nyq}} = \frac{\pi}{\Delta s}$, for which the corresponding angular frequency is the aliasing frequency of the discretization scheme, as depicted in Figure \ref{Fig:aliasing} (d).

According to the sampling theorem, spatial aliasing can be avoided by analytically spatial-bandlimiting the driving functions to the Nyquist wavenumber.
This can be performed by either spatially filtering the driving functions with an appropriate spatial low-pass filter prior to numerical evaluation, or performing bandlimitation directly in the wavenumber (modal) domain analytically.
The latter approach is implemented by \cite{hahn2022cylindrical}, and will serve as a reference solution in the following investigation.
However, the modal bandlimitation solution is available only for plane wave and point source virtual source models, at the cost of increased computational complexity.

In the following an alternative, approximate solution is presented, allowing spatial antialiasing filtering for an arbitrary virtual source model and arbitrary SSD contour, performed by FIR (finite impulse response) filtering in the temporal domain.
The technique relies on an analytic transform of spatially defined antialiasing filters to an equivalent temporal domain filter bank.

@@

\subsection{Spatial to temporal filter transform}
First, a general analytical approach is introduced allowing the transform of an arbitrary spatial filter to an equivalent temporal filter set.
From herein, subscript $x$ and $t$ distinguishes whether a filter is defined as a spatial or a temporal filter, respectively.

Assume an arbitrary filter impulse response defined in the spatial domain, denoted by $h_x(s)$.
Again, $s$ is a suitable parametrization of the SSD contour.
The spatially filtered driving function is obtained from the convolution of the conventional driving functions and the filter impulse response along the SSD contour
\begin{equation}
    d'_x(s,t) = h_x(s) \ast_s d(s,t) = \int h_x(s-s_0) \, d(s_0,t) \, \td s_0,
    \label{eq:spatial_conv}
\end{equation}
where $\ast_x$ denotes a circular convolution for a convex SSD contour or a linear convolution in case of an infinite, linear SSD.

@@

\subsubsection{Exact transformation}
The following transformation is allowed by the spatio-temporal structures of the driving functions \eqref{eq:d_wfs_td}, in which the spatial and temporal dimensions are interconnected in the argument of a Dirac-delta, due to the wave propagation characteristics.
The main idea of the derivation is that by exploiting the sifting property of the Dirac-delta in the driving functions, both spatial and temporal convolutions \eqref{eq:spatial_conv} and \eqref{eq:temp_conv} can be evaluated, from which the temporal filter impulse responses can be expressed as the function of the spatial ones.
Again, in the following the pre-equalization filter is omitted from discussion, since equalization filtering can be performed following the antialiasing process.

The temporal convolution in \eqref{eq:temp_conv} can be evaluated by substituting the general WFS driving function \eqref{eq:d_wfs_td} and exploiting the sifting property of the Dirac-delta, yielding
\begin{align}
    d'_t(s,t) = A^D(s) \, h_t \!\left(s,t- \tau(s) \right) \\
    d'_t(s,t + \tau(s) ) = A^D(s) \, h_t \!\left(s,t\right),
    \label{eq:temp_conv_eval}
\end{align}
with the second equation obtained by applying the time shift of $\tau(s)$ for the sake of later simplicity.

The shifted spatially filtered driving function reads as
\begin{multline}
    d'_x(s,t+ \tau(s) ) = \int h_x(s-s_0) \, d(s_0,t+ \tau(s) ) \, \td s_0 = \\
    \int h_x(s-s_0) \, A^D(s_0)\, \delta(s_0,t+ \left(\tau(s)-\tau(s_0)\right)  ) \, \td s_0.
    \label{eq:spat_conv2}
\end{multline}
In order to evaluate the spatial convolution the generalized sifting property of the Dirac-delta may be applied \cite{Firtha2019phd, Jackson1999}, which states that
\begin{equation}
    \int f(s_0) \delta(g(s_0)) \td s_0 = \sum_i \frac{f(s_i)}{| \frac{\partial}{\partial s_0} g(s_0) |_{s = s_i}}, \hspace{3mm} g(s_i) = 0.
\end{equation}
In the present problem the zeros of the Dirac's argument are found where
\begin{equation}
    \tau(s_i) = t + \tau(s)
    \label{eq:zero}
\end{equation}
is satisfied.
In the following it is assumed that a single zero $s_i$ exists, satisfying \eqref{eq:zero}.
This assumption means that the wavefront arrives at each SSD element at a unique time instant, being strictly true for a virtual plane wave and a linear SSD.
The derivative of the phase function is given by
\begin{equation}
    \frac{\partial}{\partial s_0} g(s_0) = - \frac{\partial}{\partial s_0} \tau(s_0) = -\frac{\hat{k}_t^P(s_0)}{c}.
\end{equation}
Here $\hat{k}_t^P(s_0)$ is the tangential component of the local propagation vector, defined as
\begin{equation}
    \hat{k}_t^P(s_0) = \mathbf{t}^{\mathrm{T}}(s_0) \hat{\vk}^P(s_0),
\end{equation}
with $\mathbf{t}^{\mathrm{T}}(s_0)$ denoting the unity long tangential vector of the SSD element at position $s_0$.

With the above assumptions the integral \eqref{eq:spat_conv2} can be evaluated, yielding
\begin{equation}
    d'_x(s,t+ \tau(s) ) =  \frac{ c }{|\hat{k}_t^P(s_i)|} \, h_x(s-s_i) \, A^D(s_i)
    .
    \label{eq:spat_conv_eval}
\end{equation}

@@

Equation \eqref{eq:temporal_transform} can be directly expressed in the spectral domain by taking the Fourier-transform of both sides, relating the frequency response of the spatial and the temporal filters.
By denoting $\mathcal{F}\left( h_x(s)\right) = H_x(k_s)$ the corresponding transform is given by
\begin{equation}
    H_t\left(s,\omega\right) = H_x\left(-\frac{\omega}{c}\hat{k}_t^P(s)\right) = H_x\left(-k_t^P(s)\right),
    \label{eq:frequency_transform}
\end{equation}
where $k_t^P(s)$ is the tangential component of the local wavenumber vector \footnote{The local wavenumber vector can be defined for a steady state sound field, pointing in the local propagation direction with the length being $\omega/c$ \cite{Firtha2019phd}}.
Hence, as the main result of the present discussion, the temporal filter transfer can be obtained from the wavenumber (modal) content of the spatial filter by simple rescaling in terms of the local wavenumber vector.

@@

\subsubsection{Application for Local Wave Field Synthesis}
The findings of the previous subsections can be straightforwardly used in order to apply the above theory directly for LWFS.
In this case a reference position $\vxref$ is prescribed inside the SSD, at which aliasing-free, amplitude and phase correct synthesis has to be ensured.
